Symmetric Circuits for Rank Gates

### 2018-03-04 sections/transparency-and-unique-extensions.tex
I have made changes to the section "symmetric circuits" and updated every other section in line with changes made.

We now show that taking the quotient of a circuit preserves important
properties, including the function computed by the circuit, whether the circuit
is symmetric, and has unique unique labels. We also show that the quotient of a
circuit has trivial syntactic-equivalence classes, and hence unique extensions.

\begin{lem}
  Let $C := \langle G, \Omega, \Sigma, \Lambda, L\rangle$ be a $(\BB,
  \rho)$-circuit of order $n$ and $C_{\equiv} = \langle G_\equiv ,
  \Omega_\equiv, \Sigma_\equiv , \Lambda_\equiv, L_\equiv \rangle$ be a quotient
  of $C$. Then $C$ and $C_{\equiv}$ compute the same function and each
  syntacitic-equivalence class in $C_\equiv$ is a singleton. Moreover, if $C$ is
  symmetric then $C_{\equiv}$ is symmetric, and for all $g \in G$, $g$ has
  unique labels in $C$ if, and only if, $[g]$ has unique labels in $C_{\equiv}$.
  \label{lem:quotient-circuits-preserve}
\end{lem}
\begin{proof}
  We first prove that $C_\equiv$ and $C$ compute the same function. Let
  $\mathcal{A}$ be a $\rho$-structure of size $n$ and let $\gamma$ be a
  bijection from $U$ to $[n]$. We now show that for all $g \in G$, $C_\equiv
  [\gamma \mathcal{A}]([g]) = C[\gamma \mathcal{A}](g)$. We do this by induction
  on the depth of a gate. Suppose $g \in G$ has depth $0$. In this case $g$ is
  an input gate and the result follows trivially. Suppose $g$ is an internal
  gate and suppose for all $h$ of depth less than $g$ we have that
  $C_\equiv[\gamma \mathcal{A}]([h]) = C[\gamma \mathcal{A}](h)$. We have that
  there exists $g' \in [g]$ such that $L_\equiv ([g]) = L(g') /_\equiv$ and
  $\Sigma_\equiv([g]) = \Sigma (g) = \Sigma (g')$. We have from
  Lemma~\ref{lem:syntactic-equivalence-equal-function} and the inductive
  hypothesis that there exists $\lambda \in \aut(g)$ such that $L_\equiv ([g]) =
  L(g') = L(g) \lambda$. It follows from the fact that $\Sigma(g)$ is a
  structured function that $C_\equiv[\gamma \mathcal{A}]([g]) = \Sigma_\equiv
  ([g])(L^{\gamma \mathcal{A}}_\equiv([g])) = \Sigma (g') (L^{\gamma
    \mathcal{A}}(g')) = \Sigma (g)(L^{\gamma \mathcal{A}}(g) \lambda) = \Sigma
  (g) (L^{\gamma \mathcal{A}}(g)) = C[\gamma \mathcal{A}](g)$.

  We now show that each syntactic-equivalence class in $C_\equiv$ is a
  singleton. Suppose $[g], [h] \in G_\equiv$ and suppose $[g] \equiv [h]$. If
  $[g]$ and $[h]$ are both input or output gates then $[g] = [h]$. Suppose $[g]$
  and $[h]$ are internal gates that are not output gates. Then $\Sigma (g) =
  \Sigma_\equiv ([g]) = \Sigma_\equiv([h]) = \Sigma(h)$. There exists $g', h'
  \in G$ such that $g' \equiv g$ and $h' \equiv h$, and $L_\equiv ([g]) = L(g')
  /_\equiv$ and $L_\equiv ([h]) = L(h') /_\equiv$. It follows from $[g] \equiv
  [h]$ that $L(g') /_\equiv$ is isomorphism-equivalent to $L(h') /_\equiv$. It
  follows that $g' \equiv h'$, and so $[g] = [h]$.

@@

  We now define the circuit $C_1$ formally. Let $C_1 := \langle G_1, \Omega_1,
  \Sigma_1 , \Lambda_1, L_1 \rangle$ be defined as follows. Let $G_1 = G_0
  \uplus \{ g_\lor\}$. Let $\Omega_1 = \Omega_0$ and $\Lambda_1 = \Lambda_0$.
  Let $g \in G_1$. If $g = g_\lor$ then $\Sigma_1 (g) = \lor[2]$, $L_1(g)(1) =
  g_0$, and $L_1(g)(2) = g_1$. If $g \in G_0$, $\Sigma(g) = \lor[2]$ and $H_g =
  \{g_0, g_1\}$ then $\Sigma_1 (g) = \lor[3]$, $L_1(g)(1) = L_0(g)(1)$,
  $L_1(g)(2) = L_0(g)(2)$, and $L_1(g)(3) = g_\lor$. If $g \in G_0$ and $g =
  \land[k]$ for some $k \in \nats$ then $\Sigma_1(g) = \land[k+1]$ and
  $L_1(g)(i) = L_0(g)(i)$ for all $i \in [k]$ and $L_1(g)(k+1) = g_\lor$.
  Otherwise let $\Sigma_1(g) = \Sigma_0(g)$ and $L_1(g) = L_0(g)$.

@@

  We have from Lemma~\ref{lem:quotient-circuits-preserve} that $C_0$ has trivial
  syntactic-equivalence classes. Since $g_\lor$ is the only two input $\OR$-gate
  with exactly the two constant gates as children, it follows that $g_\lor$ is
  contained in a singleton syntactic-equivalence class. It can be shown by
  induction that if $g, g' \in G_0$ are syntactically equivalent in $C_1$ then
  they must be syntactically-equivalent in $C_0$. It follows that if $g$ and
  $g'$ are syntactically-equivalent in $C_1$ then $g = g'$. We thus have that
  $C_1$ has trivial syntactic-equivalence classes.

  Let $C' := \{G', \Omega', \Sigma' \Lambda' L'\}$ be defined as follows. For
  each $h \in G_1$ let $c^h_g := \vert L_1(g)^{-1}(h) \vert$ if $h \in H_g$ and
  $c^h_g := 0$ otherwise. For each $h \in G_1$ let $c^h$ be the maximum of
  $c^h_g$ for $g \in G_1$. For each $h \in G_1$ if $c^h > 1$ let $G_h :=
  \{g^{h}_{1}, ldots, g^h_{c_h-1}\}$ otherwise let $G_h := \emptyset$. Let
  $G_{\land} := \biguplus_{h \in G_1} G_h$ and $G' = G_{\land} \uplus G_1$. Let
  $\Omega' = \Omega_1$ and $\Lambda' = \Lambda_1$. For $g \in G'$ if $g \in G_1$
  let $\Sigma'(g) = \Sigma_1(g)$ and otherwise let $\Sigma'(g) = \land[1]$. For
  each $g \in G_1$ and $h \in H_g$ let $\{x^{h,g}_0, x^{h,g}_1 ,\ldots,
  x^{h,g}_{c^h_g - 1}\} := L_1(g)^{-1}(h)$ be a fixed ordering of the set
  $L_1(g)^{-1}(h)$. Let $g \in G'$. If $g \in G_1$ and $x \in \ind(g)$ then
  there exists $h \in H_g$ and $i \in \{0, \ldots , c^h_g -1\}$ such that $x =
  x^{h,g}_i$, and we let $L'(g)(x) = h$ if $i = 0$ and $L'(g)(x^{h,g}_i) =
  g^h_i$ otherwise. If $g \in G_{\land}$ then $g = g^h_i$ for some $h \in G_1$
  and $i \in [c^h_g]$, and we let $L'(g)(1) = h$ if $i = 1$ and $L'(g)(1) =
  g^h_{i-1}$ otherwise.

@@

  \begin{claim}
    $C'$ has trivial syntactic-equivalence classes.
  \end{claim}
  \begin{proof}
    We prove this result by induction on depth. Let $g \in G'$. If $g$ has depth
    $0$ then $g$ is an input gate and so $[g] = \{g\}$. Let $g \in G'$ be an
    internal gate in $C'$ and suppose for each gate $h$ of depth less than $g$
    we have that $[h] = \{ h \}$. Let $g' \in G'$ and suppose $g \equiv' g'$. We
    now show that $g = g'$, and so prove the claim. Since $g \equiv' g'$ it
    follows that $\Sigma (g) = \Sigma(g')$ and both $g$ and $g'$ have the same
    depth. Moreover, from the inductive hypothesis, we have that there exists
    $\lambda \in \aut(g)$ such that for all $x \in \ind(g)$, $L'(g) (x) =
    L'(g')(x)$. It follows that $H_{g}' = H_{g'}'$. Moreover, from the inductive
    hypothesis the children of $g$ and $g'$ must be contained in singleton
    syntactic-equivalence classes, and so , since $C'$ has injective labels, it
    follows that $g$ and $g'$ have unique labels in $C'$.

@@

    Suppose $g \in G_{\land}$. Then there exists $h \in G_1$ and $i \in [c^h -
    1]$ such that $g = g^h_i$. Suppose $g' \in G_1$. Then $\Sigma'(g) =
    \land[1]$. But, from the construction of $C_1$, there are no single-input
    $\AND$-gates in $G_1$. It follows $\Sigma'(g') \neq \Sigma'(g)$ and so $g'
    \not\equiv' g$, a contradiction, and so we must have $g' \in G_{\land}$. If
    $i = 1$ then $\{ h \} = H_g' = H_{g'}'$. Since the only gate in $G_\land$
    that has $h$ as a child gate is $g^h_1$, it follows that $g = g^h_1 = g'$.
    If $i > 1$ then $\{g^{h}_{i-1}\} = H_g' = H_{g'}'$. Since $g^h_i$ is the
    only gate in $G_\land$ that has $g^h_{i-1}$ as a child gate, we have $g =
    g^h_i = g'$. It follows that if $g \in G_\land$ then $g = g'$.

    Suppose $g \in G_1$. We have already shown that if $g' \in G_\land G_1$ then
    $g \not\equiv' g'$, a contradiction, and so we must have $g' \in G_1$. Since
    $g \equiv' g'$ there exists $\lambda \in \aut(g)$ such that for all $x \in
    \ind(g)$, $L'(g)(\lambda x) = L'(g')(x)$. From the construction, we have
    that for all $x \in \ind(g)$ and $h \in G_1$, $L'(g)(x) \in \{h, g^h_1 ,
    \ldots, g^h_{c^h_g -1}\}$ if, and only if, $L_1(g)(x) = h$. Suppose $g$ is
    not a symmetric gate. Since $C$ is transparent, $C_0$ is transparent and so
    $C_1$ is transparent. It follows that for all $h \in H_g$, $H_g = H_{g'}$,
    and $c^h_g = 1 = c^h_{g'}$. It follows that for all $x \in \ind(g)$, there
    exists a unique $h \in G_1$ such that $L_1(g)(\lambda x) = L'(g)(\lambda x)
    = L'(g')(x) = L_1 (g')(x) = h$. It follows that $g \equiv g'$ and so $g =
    g'$. Suppose $g$ is a symmetric gate. Let $x \ind(g)$ and $h = L_1(g)(x)$.
    If $L'(g)(\lambda x) = L'(g')(x)= h$ then $L_1(g)(\lambda x) = h$ if, and
    only if, $L'(g)(\lambda x ) = h* = L'(g')(x)$, where $h* = h$ or $h* =
    g^{h}_i$, for some $i \in [c^h_g -1]$, if, and only if, $L_1(g) (x) = h$. It
    follows that for all $x \in \ind(g)$, $L_1 (g) (\lambda x) = L_1 (g)(x)$,
    and so $g \equiv g'$ and so $g = g'$.

@@

  Since $C'$ has injective labels and trivial syntactic-equivalence classes it
  follows $C'$ has unique labels.

@@

\begin{lem}
  There is an algorithm takes as input a $(\mathbb{B}, \rho)$-circuit $C$ of
  order $n$ with unique labels and $\sigma \in \sym_n$ and outputs for each gate
  $g$ its image under the unique automorphism extending $\sigma$ (if it exists).
  This algorithm runs in time polynomial in the combined size of the input
  circuit and the encoding of the permutation.
  \label{lem:compute-automorphisms}
\end{lem}
\begin{proof}
  Let $C := \langle G, \Omega, \Sigma, \Lambda, L \rangle$. We recursively build
  up the mapping $\pi$ extending $\sigma$. If at some point in the recursive
  construction we arrive at a point where no mapping for $g$ can be found we
  halt at that point and return that no automorphism exists.

@@

We now present a reduction from the graph-isomorphism problem to the problem of
deciding if two gates in a circuit are syntactically-equivalent. In fact, we
prove a stronger result, presenting a reduction from the graph-isomorphism
problem to the problem of computing the syntactic-equivalence relation over an
ostensibly simpler class of circuits.

\begin{prop}
  Let $\rho$ be any non-empty relational vocabulary and let $n, r, p \in \nats$,
  with $p$ prime. There is a a polynomial time reduction from the
  graph-isomorphism problem to the problem of deciding if a given pair of gates
  in a given symmetric rank-circuit for structures of size $n$ that (i) has
  injective labels, (ii) contains no constant gates, and (iii) contains at most
  two rank gates with bound $r$ and prime $p$, are syntactically equivalent.
  \label{prop:syntactic-graph-iso}
\end{prop}
\begin{proof}
  We have from \cite{} that there is a polynomial time reduction from the graph
  isomorphism problem to the bipartite isomorphism problem. The bipartite
  isomorphism problem is the decision problem for the set of all pairs of
  partitioned bipartite graphs $B_1 := (U_1, V_1, E_1)$ and $B_2 := (U_2, V_2,
  E_2)$ such that there exists a graph isomorphism $\pi : B_1 \rightarrow B_2$
  with $\restr{\pi}{U_1} = U_2$ and $\restr{\pi}{V_1} = V_2$.

@@

  Let $R$ be a relation symbol in $\rho$ and let $k := \arty(R) > 0$. Let $G_{R}
  := \{g_{R, \vec{c}} : \vec{c} \in [n]^k\}$, $G_{\text{mid}} := \{g_{\lor},
  g_{\land}, g_{\text{out}} \}$, $G_{\rank} : = \{ g^1_{\rank}, g^2_{\rank}\}$,
  and $G_{\text{nodes}} := \{g_{i, (u,v)} : i \in [2] \, (u,v) \in [a] \times
  [b] \}$. Let $C = \langle G, \Omega, \Sigma, \Lambda, L \rangle$ be a
  $(\{\rank^{1}_2\}, \rho)$-circuit of order $n$ defined as follows. Let $G =
  G_R \cup G_{\text{mid}} \cup G_{\rank} \cup G_{\text{nodes}}$ and $\Omega$ be
  the $0$-ary function $\land_{\text{out}}$. For each $\vec{c} \in
  [n]^{\arty(R)}$ let $\Lambda_R(\vec{c}) = g_{R, \vec{c}}$. Define $\Sigma$ as
  follows. For each $g \in G$,
  \begin{itemizens}
    \setlength\itemsep{0mm}
  \item if $g = \land_{\text{out}}$ then $\Sigma(g) = \land[2]$,
  \item if $g \in G_\rank$ then $\Sigma(g) = \rank^r_p [a,b]$,
  \item if $g \in G_{\text{nodes}}$ then $\Sigma(g) = \land[1]$,
  \item if $g = \land[n^k]$, $\Sigma (g) = \land[n^k]$ and if $g = g_{\lor}$
    then $\Sigma(g) = \lor[n^k]$, and
  \item if $g \in G_R$ then $\Sigma(g) = R$.
  \end{itemizens}
  Define $L$ as follows. For each $g \in G$,
  \begin{itemizens}
    \setlength\itemsep{0mm}
  \item if $g = \land_{\text{out}}$ then for each $i \in [2]$, $L(g)(i) :=
    g^i_{\rank}$,
  \item if $g \in G_\rank$ and $g = g^i_{\rank}$ then for all $(p,q) \in
    \ind(g)$, $L(g)(p,q) = g_{i, (p,q)}$,
  \item if $g \in G_{\text{nodes}}$ and $g = g_{i, (p,q)}$ then if $(p, q \in
    E_i)$ then $L(g)(1) = g_{\land}$, otherwise $L(g)(1) = g_\lor$, and
  \item if $g = g_\land$ or $g = g_\lor$ then for all $p \in [n^k]$ we have that
    $L(g)(p)$ equals $\Lambda^{-1}(\vec{c}_p)$, where $\vec{c}_p$ is the $p$th
    element of $[n]^k$ in the lexicographical ordering on $[n]^k$.
  \end{itemizens}

@@

We now show that there is a reduction from the graph-isomorphism problem to the
problem of deciding if a given circuit is symmetric. In fact, we prove a
stronger result, showing that this reduction holds even if we restrict ourselves
to the class of circuits with trivial syntactic-equivalence classes and in
which, all but two gates in the circuit have unique labels.

\begin{prop}
  The graph-isomorphism problem is polynomial-time reducible to the problem of
  deciding if a circuit with trivial syntactic-equivalence classes and in which
  all but two gates in the circuit have unique labels is symmetric.
  \label{prop:graph-iso-symmetric}
\end{prop}
\begin{proof}
  We use a similar approach as in Proposition~\ref{prop:syntactic-graph-iso},
  and so proving a reduction from the bipartite-graph-isomorphism to the problem
  of deciding if a circuit is symmetric.

@@

Note that, in Proposition~\ref{prop:graph-iso-symmetric} we again have a
hardness result for deciding a basic circuit property even if we restrict
ourselves to circuits with trivial syntactic-equivalence classes (and so ones
where each gate has unique children). It is possible to alter this reduction,
using a construction analogous to the one used in the proof of
Proposition~\ref{prop:syntactic-graph-iso}, to prove that deciding symmetry is
also hard for circuits with injective labels. In contrast, we have from
Proposition~\ref{lem:computing-support-orbit-index} that we can decide symmetry
for transparent circuits (i.e. circuits where every non-symmetric gate has
unique children and injective labels) in polynomial-time. These observations
again suggest both the robustness of this hardness result, and the importance of
the transparency condition.

>>>>

We now show that taking the quotient of a circuit preserves important
properties, including the function computed by the circuit, whether the circuit
is symmetric, and has unique unique labels. We also show that the quotient of a
circuit is reduced, and hence has unique extensions.

\begin{lem}
  Let $C := \langle G, \Omega, \Sigma, \Lambda, L\rangle$ be a $(\BB,
  \rho)$-circuit of order $n$ and $C_{\equiv} = \langle G_\equiv ,
  \Omega_\equiv, \Sigma_\equiv , \Lambda_\equiv, L_\equiv \rangle$ be a quotient
  of $C$. Then $C_\equiv$ is reduced and $C$ and $C_{\equiv}$ compute the same
  function. Moreover, if $C$ is symmetric then $C_{\equiv}$ is symmetric, and
  for all $g \in G$, $g$ has unique labels in $C$ if, and only if, $[g]$ has
  unique labels in $C_{\equiv}$.
  \label{lem:quotient-circuits-preserve}
\end{lem}
\begin{proof}
  We first prove that $C_\equiv$ and $C$ compute the same function. Let
  $\mathcal{A}$ be a $\rho$-structure of size $n$ and let $\gamma$ be a
  bijection from $U$ to $[n]$. We now show that for all $g \in G$, $C_\equiv
  [\gamma \mathcal{A}]([g]) = C[\gamma \mathcal{A}](g)$. We do this by induction
  on the depth of a gate. Suppose $g \in G$ has depth $0$. In this case $g$ is
  an input gate and the result follows trivially. Suppose $g$ is an internal
  gate and suppose for all $h$ of depth less than $g$ we have that
  $C_\equiv[\gamma \mathcal{A}]([h]) = C[\gamma \mathcal{A}](h)$. We have that
  there exists $g' \in [g]$ such that $L_\equiv ([g]) = L(g') /_\equiv$ and
  $\Sigma_\equiv([g]) = \Sigma (g) = \Sigma (g')$. We have from
  Lemma~\ref{lem:syntactic-equivalence-equal-function} and the inductive
  hypothesis that there exists $\lambda \in \aut(g)$ such that $L_\equiv ([g]) =
  L(g') = L(g) \lambda$. It follows from the fact that $\Sigma(g)$ is a
  structured function that $C_\equiv[\gamma \mathcal{A}]([g]) = \Sigma_\equiv
  ([g])(L^{\gamma \mathcal{A}}_\equiv([g])) = \Sigma (g') (L^{\gamma
    \mathcal{A}}(g')) = \Sigma (g)(L^{\gamma \mathcal{A}}(g) \lambda) = \Sigma
  (g) (L^{\gamma \mathcal{A}}(g)) = C[\gamma \mathcal{A}](g)$.

  We now show that $C_\equiv$ is reduced. Suppose $[g], [h] \in G_\equiv$ and
  suppose $[g] \equiv [h]$. If $[g]$ and $[h]$ are both input or output gates
  then $[g] = [h]$. Suppose $[g]$ and $[h]$ are internal gates that are not
  output gates. Then $\Sigma (g) = \Sigma_\equiv ([g]) = \Sigma_\equiv([h]) =
  \Sigma(h)$. There exists $g', h' \in G$ such that $g' \equiv g$ and $h' \equiv
  h$, and $L_\equiv ([g]) = L(g') /_\equiv$ and $L_\equiv ([h]) = L(h')
  /_\equiv$. It follows from $[g] \equiv [h]$ that $L(g') /_\equiv$ is
  isomorphism-equivalent to $L(h') /_\equiv$. It follows that $g' \equiv h'$,
  and so $[g] = [h]$.

@@

  We now define the circuit $C_1$ formally. Let $C_1 := \langle G_1, \Omega_1,
  \Sigma_1 , \Lambda_1, L_1 \rangle$ be defined as follows. Let $G_1 = G_0
  \uplus \{ g_\lor\}$. Let $\Omega_1 = \Omega_0$ and $\Lambda_1 = \Lambda_0$.
  Let $g \in G_1$. If $g = g_\lor$ then $\Sigma_1 (g) = \OR[2]$, $L_1(g)(1) =
  g_0$, and $L_1(g)(2) = g_1$. If $g \in G_0$, $\Sigma(g) = \OR[2]$ and $H_g =
  \{g_0, g_1\}$ then $\Sigma_1 (g) = \OR[3]$, $L_1(g)(1) = L_0(g)(1)$,
  $L_1(g)(2) = L_0(g)(2)$, and $L_1(g)(3) = g_\lor$. If $g \in G_0$ and $g =
  \AND[k]$ for some $k \in \nats$ then $\Sigma_1(g) = \AND[k+1]$ and $L_1(g)(i)
  = L_0(g)(i)$ for all $i \in [k]$ and $L_1(g)(k+1) = g_\lor$. Otherwise let
  $\Sigma_1(g) = \Sigma_0(g)$ and $L_1(g) = L_0(g)$.

@@

  We have from Lemma~\ref{lem:quotient-circuits-preserve} that $C_0$ is reduced.
  Since $g_\lor$ is the only two input $\OR$-gate with exactly the two constant
  gates as children, it follows that $g_\lor$ is contained in a singleton
  syntactic-equivalence class. It can be shown by induction that if $g, g' \in
  G_0$ are syntactically equivalent in $C_1$ then they must be
  syntactically-equivalent in $C_0$. It follows that if $g$ and $g'$ are
  syntactically-equivalent in $C_1$ then $g = g'$. We thus have that $C_1$ is
  reduced.

  Let $C' := \{G', \Omega', \Sigma' \Lambda' L'\}$ be defined as follows. For
  each $h \in G_1$ let $c^h_g := \vert L_1(g)^{-1}(h) \vert$ if $h \in H_g$ and
  $c^h_g := 0$ otherwise. For each $h \in G_1$ let $c^h$ be the maximum of
  $c^h_g$ for $g \in G_1$. For each $h \in G_1$ if $c^h > 1$ let $G_h :=
  \{g^{h}_{1}, ldots, g^h_{c_h-1}\}$ otherwise let $G_h := \emptyset$. Let
  $G_{\land} := \biguplus_{h \in G_1} G_h$ and $G' = G_{\land} \uplus G_1$. Let
  $\Omega' = \Omega_1$ and $\Lambda' = \Lambda_1$. For $g \in G'$ if $g \in G_1$
  let $\Sigma'(g) = \Sigma_1(g)$ and otherwise let $\Sigma'(g) = \AND[1]$. For
  each $g \in G_1$ and $h \in H_g$ let $\{x^{h,g}_0, x^{h,g}_1 ,\ldots,
  x^{h,g}_{c^h_g - 1}\} := L_1(g)^{-1}(h)$ be a fixed ordering of the set
  $L_1(g)^{-1}(h)$. Let $g \in G'$. If $g \in G_1$ and $x \in \ind(g)$ then
  there exists $h \in H_g$ and $i \in \{0, \ldots , c^h_g -1\}$ such that $x =
  x^{h,g}_i$, and we let $L'(g)(x) = h$ if $i = 0$ and $L'(g)(x^{h,g}_i) =
  g^h_i$ otherwise. If $g \in G_{\land}$ then $g = g^h_i$ for some $h \in G_1$
  and $i \in [c^h_g]$, and we let $L'(g)(1) = h$ if $i = 1$ and $L'(g)(1) =
  g^h_{i-1}$ otherwise.

@@

  \begin{claim}
    $C'$ is reduced
  \end{claim}
  \begin{proof}
    We prove this result by induction on depth. Let $g \in G'$. If $g$ has depth
    $0$ then $g$ is an input gate and so $[g] = \{g\}$. Let $g \in G'$ be an
    internal gate in $C'$ and suppose for each gate $h$ of depth less than $g$
    we have that $[h] = \{ h \}$. Let $g' \in G'$ and suppose $g \equiv' g'$. We
    now show that $g = g'$, and so prove the claim. Since $g \equiv' g'$ it
    follows that $\Sigma (g) = \Sigma(g')$ and both $g$ and $g'$ have the same
    depth. Moreover, from the inductive hypothesis, we have that there exists
    $\lambda \in \aut(g)$ such that for all $x \in \ind(g)$, $L'(g) (x) =
    L'(g')(x)$. It follows that $H_{g}' = H_{g'}'$. Moreover, from the inductive
    hypothesis the children of $g$ and $g'$ must be contained in a singleton
    syntactic-equivalence classes, and so, since $C'$ has injective labels, it
    follows that $g$ and $g'$ have unique labels in $C'$.

@@

    Suppose $g \in G_{\land}$. Then there exists $h \in G_1$ and $i \in [c^h -
    1]$ such that $g = g^h_i$. Suppose $g' \in G_1$. Then $\Sigma'(g) =
    \AND[1]$. But, from the construction of $C_1$, there are no single-input
    $\AND$-gates in $G_1$. It follows $\Sigma'(g') \neq \Sigma'(g)$ and so $g'
    \not\equiv' g$, a contradiction, and so we must have $g' \in G_{\land}$. If
    $i = 1$ then $\{ h \} = H_g' = H_{g'}'$. Since the only gate in $G_\land$
    that has $h$ as a child gate is $g^h_1$, it follows that $g = g^h_1 = g'$.
    If $i > 1$ then $\{g^{h}_{i-1}\} = H_g' = H_{g'}'$. Since $g^h_i$ is the
    only gate in $G_\land$ that has $g^h_{i-1}$ as a child gate, we have $g =
    g^h_i = g'$. It follows that if $g \in G_\land$ then $g = g'$.

    Suppose $g \in G_1$. We have already shown that if $g' \in G_\land$ then $g
    \not\equiv' g'$, a contradiction, and so we must have $g' \in G_1$. Since $g
    \equiv' g'$ there exists $\lambda \in \aut(g)$ such that for all $x \in
    \ind(g)$, $L'(g)(\lambda x) = L'(g')(x)$. From the construction, we have
    that for all $x \in \ind(g)$ and $h \in G_1$, $L'(g)(x) \in \{h, g^h_1 ,
    \ldots, g^h_{c^h_g -1}\}$ if, and only if, $L_1(g)(x) = h$. Suppose $g$ is
    not a symmetric gate. Since $C$ is transparent, $C_0$ is transparent and so
    $C_1$ is transparent. It follows that for all $h \in H_g$, $H_g = H_{g'}$,
    and $c^h_g = 1 = c^h_{g'}$. It follows that for all $x \in \ind(g)$, there
    exists a unique $h \in G_1$ such that $L_1(g)(\lambda x) = L'(g)(\lambda x)
    = L'(g')(x) = L_1 (g')(x) = h$. It follows that $g \equiv g'$ and so $g =
    g'$. Suppose $g$ is a symmetric gate. Let $x \ind(g)$ and $h = L_1(g)(x)$.
    If $L'(g)(\lambda x) = L'(g')(x)= h$ then $L_1(g)(\lambda x) = h$ if, and
    only if, $L'(g)(\lambda x ) = h* = L'(g')(x)$, where $h* = h$ or $h* =
    g^{h}_i$, for some $i \in [c^h_g -1]$, if, and only if, $L_1(g) (x) = h$. It
    follows that for all $x \in \ind(g)$, $L_1 (g) (\lambda x) = L_1 (g)(x)$,
    and so $g \equiv g'$ and so $g = g'$.

@@

  Since $C'$ has injective labels and is reduced it follows $C'$ has unique
  labels.

@@

\begin{lem}
  There is an algorithm takes as input a $(\BB, \rho)$-circuit $C$ of order $n$
  with unique labels and $\sigma \in \sym_n$ and outputs for each gate $g$ its
  image under the unique automorphism extending $\sigma$ (if it exists). This
  algorithm runs in time polynomial in the combined size of the input circuit
  and the encoding of the permutation.
  \label{lem:compute-automorphisms}
\end{lem}
\begin{proof}
  Let $C := \langle G, \Omega, \Sigma, \Lambda, L \rangle$. We recursively build
  up the mapping $\pi$ extending $\sigma$. If at some point in the recursive
  construction we arrive at a point where no mapping for $g$ can be found we
  halt at that point and return that no automorphism exists.

@@

We now present a reduction from the graph-isomorphism problem to the problem of
deciding if two gates in a circuit are syntactically-equivalent. In fact, we
prove a stronger result, presenting a reduction from the graph-isomorphism
problem to the problem of computing the syntactic-equivalence relation over a
more constrained class of circuits.

\begin{prop}
  Let $\rho$ be any non-empty relational vocabulary and let $n, r, p \in \nats$,
  with $p$ prime. There is a a polynomial time reduction from the
  graph-isomorphism problem to the problem of deciding if a given pair of gates
  in a given symmetric rank-circuit for structures of size $n$ that (i) has
  injective labels, (ii) contains no constant gates, and (iii) contains at most
  two rank gates with bound $r$ and prime $p$, are syntactically-equivalent.
  \label{prop:syntactic-graph-iso}
\end{prop}
\begin{proof}
  We have from \cite{} that there is a polynomial time reduction from the graph
  isomorphism problem to the bipartite isomorphism problem. The bipartite
  isomorphism problem is the decision problem for the set of all pairs of
  partitioned bipartite graphs $B_1 := (U_1, V_1, E_1)$ and $B_2 := (U_2, V_2,
  E_2)$ such that there exists a graph isomorphism $\pi : B_1 \rightarrow B_2$
  with $\restr{\pi}{U_1} = U_2$ and $\restr{\pi}{V_1} = V_2$.

@@

  Let $R$ be a relation symbol in $\rho$ and let $k := \arty(R) > 0$. Let $G_{R}
  := \{g_{R, \vec{c}} : \vec{c} \in [n]^k\}$, $G_{\text{mid}} := \{g_{\lor},
  g_{\land}, g_{\text{out}} \}$, $G_{\rank} : = \{ g^1_{\rank}, g^2_{\rank}\}$,
  and $G_{\text{nodes}} := \{g_{i, (u,v)} : i \in [2] \, (u,v) \in [a] \times
  [b] \}$. Let $C = \langle G, \Omega, \Sigma, \Lambda, L \rangle$ be a
  $(\{\rank^{1}_2\}, \rho)$-circuit of order $n$ defined as follows. Let $G =
  G_R \cup G_{\text{mid}} \cup G_{\rank} \cup G_{\text{nodes}}$ and $\Omega$ be
  the $0$-ary function $g_{\text{out}}$. For each $\vec{c} \in
  [n]^{\arty(R)}$ let $\Lambda_R(\vec{c}) = g_{R, \vec{c}}$. Define $\Sigma$ as
  follows. For each $g \in G$,
  \begin{myitemize}
  \item if $g = g_{\text{out}}$ then $\Sigma(g) = \AND[2]$,
  \item if $g \in G_\rank$ then $\Sigma(g) = \RANK^r_p [a,b]$,
  \item if $g \in G_{\text{nodes}}$ then $\Sigma(g) = \AND[1]$,
  \item if $g = \AND[n^k]$, $\Sigma (g) = \AND[n^k]$ and if $g = g_{\lor}$
    then $\Sigma(g) = \OR[n^k]$, and
  \item if $g \in G_R$ then $\Sigma(g) = R$.
  \end{myitemize}
  Define $L$ as follows. For each $g \in G$,
  \begin{myitemize}
    \setlength\itemsep{0mm}
  \item if $g = g_{\text{out}}$ then for each $i \in [2]$, $L(g)(i) :=
    g^i_{\rank}$,
  \item if $g \in G_\rank$ and $g = g^i_{\rank}$ then for all $(p,q) \in
    \ind(g)$, $L(g)(p,q) = g_{i, (p,q)}$,
  \item if $g \in G_{\text{nodes}}$ and $g = g_{i, (p,q)}$ then if $(p, q \in
    E_i)$ then $L(g)(1) = g_{\land}$, otherwise $L(g)(1) = g_\lor$, and
  \item if $g = g_\land$ or $g = g_\lor$ then for all $p \in [n^k]$ we have that
    $L(g)(p)$ equals $\Lambda^{-1}(\vec{c}_p)$, where $\vec{c}_p$ is the $p$th
    element of $[n]^k$ in the lexicographical ordering on $[n]^k$.
  \end{myitemize}

@@

We now show that there is a reduction from the graph-isomorphism problem to the
problem of deciding if a given circuit is symmetric. In fact, we prove a
stronger result, showing that this reduction holds even if we restrict ourselves
to the class of reduced circuits in which all but two gates in the circuit have
injective labels. In other words, the class of circuits that have unique
children and \emph{almost} have unique labels but for two gates.

\begin{prop}
  The graph-isomorphism problem is polynomial-time reducible to the problem of
  deciding if a reduced circuit in which all but two gates in the circuit have
  unique labels is symmetric.
  \label{prop:graph-iso-symmetric}
\end{prop}
\begin{proof}
  We use a similar approach as in Proposition~\ref{prop:syntactic-graph-iso},
  and so proving a reduction from the bipartite-graph-isomorphism to the problem
  of deciding if a circuit is symmetric.

@@

Note that, in Proposition~\ref{prop:graph-iso-symmetric} we again have a
hardness result for deciding a basic circuit property even if we restrict
ourselves to reduced circuits (and so ones where each gate has unique children).
It is possible to alter this reduction, using a construction analogous to the
one used in the proof of Proposition~\ref{prop:syntactic-graph-iso}, to prove
that deciding symmetry is also hard for circuits with injective labels. In
contrast, we have from Proposition~\ref{lem:computing-support-orbit-index} that
we can decide symmetry for transparent circuits (i.e. circuits where every
non-symmetric gate has unique children and injective labels) in polynomial-time.
These observations again suggest both the robustness of this hardness result,
and the importance of the transparency condition.
